A Global Genome-wide Scan with Optimal Cutoff Mining for Emerging Biomarkers in Head and Neck Squamous Cell Carcinoma

### 2021-01-02 main.tex
done for submitting
Updates from Overleaf

The 9416 Kaplan-Meier plots with associated Cox's univariate and multivariate tables were generated at workflow step 1 (see Figure \ref{fig:figure1}) and justified by the ranking of hazard ratios.
By uncorrected P-value below 0.05, we selected 967 genes in which \acrfull{hr} is greater than 1.5 or less than 0.5 (see Figure \ref{fig:figure2}A univariate, and Figure \ref{fig:figure2}B  multivariate plots). At the final step, a Bonferroni P-value correction was used to yield the twenty candidate genes, under the stringent criteria (see Figure \ref{fig:figure2}C, D). The ten candidates, including DKK1, CAMK2N1, STC2, PGK1, SURF4, USP10, NDFIP1, FOXA2, STIP1, and DKC1, have significantly associated with the poor prognosis of \acrshort{os} (see Table \ref{table:table1}),
%\subsubsection{Figure 3}
while the other ten genes were over-expressed in the better survival patients, named as ZNF557, ZNF266, IL19, MYO1H, FCGBP, LOC148709, EVPLL, PNMA5, IQCN (previous name as KIAA1683), and NPB (see Table \ref{table:table3}, with their full name of genes).
We made a volcano plot for 9416 genes by Kaplan-Meier P-value (less than 0.05, obtained during cutoff finding procedure) against the Cox hazard ratio (see Figure \ref{fig:figure3}). The plot revealed that the most significant (Bonferroni-adjusted P-value $< 0.05$) candidate genes are located above the dotted line. At the same time, Cox's HR separated them on the two-side with prognostic impact.

%\subsubsection{Figure 2:h}
\begin{figure}[hp]
\centering
\includegraphics[width=\textwidth]{Figure2.pdf}
\bcaption{\acrshort{hnscc} Cox's hazard ratio and P-value plots.}
{(a) Univariate HR versus uncorrected P-value; (b) Multivariate HR versus uncorrected P-value; (c) Univariate HR versus Bonferroni corrected P-value; and (d) Multivariate HR versus Bonferroni corrected P-value.}
\label{fig:figure2}
\end{figure}

@@

%\subsubsection{Figure 3:h}
\begin{figure}[hp]
\centering
\includegraphics[width=\textwidth]{TCGA_HNSC_Optimal_Overall_allPlot_unKM_P_multiHR-Figure3.pdf} % .png from PvalueplotKM_20genes_Bonf.pdf
\bcaption{Volcano plot of genes under survival analyses.}
{X axis: unadjusted P-value of Kaplan-Meier survival (-log10 transformed).
Y axis: multivariate hazard ratio from Cox proportional regression model.
Dotted line: significant Bonferroni corrected P-value. 
Red circles mark 10 candidate genes, which impact on poor prognosis ($HR>=1.5$). Green circles mark 10 genes, which affect on better survival ($HR<=0.5$).}
\label{fig:figure3}
\end{figure}
% new figure of volcano plot: TCGA_HNSC_Optimal_Overall_allPlot_unKM_P_multiHR.pdf, -log10(raw KM P-value) vs Cox multivariate HR
% X axis: "frequency" (number of uncorrected P-values less than 0.05 during Kaplan-Meier cutoff finding procedure). X axis: adjusted Kaplan-Meier P-value.

@@

Our top 1 candidate is \acrfull{DKK1} (see Figure \ref{fig:figure4}A). The Kaplan-Meier curve revealed 227 patients bearing the higher expression of \acrshort{DKK1} were suffered from only 40\% of 5-year \acrshort{os} rate. In comparison, the other 187 patients with lower expression (the cutoff at -0.312(RSEM)) have been a better prognosis (adjusted P-value as $0.001$).
Figure \ref{fig:figure4}B's cumulative P-value plot showed that the uncorrected 116 P-values ($< 0.05$) were estimated by a serial cut from 125 to 290 persons for grouping the cohort in our cutoff finding procedure (cutofFinder\_func.R, see Figure \ref{fig:figure1}, "HNSCC cohort"). The smallest P-value (\num{8.9e-8}), when cut on n=187 (45.2\% of 414), was defined as optimal P-value.
Conversely, the most associated gene with better survival is \acrfull{ZNF557}. In Figure \ref{fig:figure4}C, a Kaplan-Meier curve revealed 264 patients bearing the higher expression of \acrshort{ZNF557} have 55\% of 5-year OS survival rate (adjusted P-value = $0.001$). The cutoff finding procedure (cutofFinder\_func.R) generated cumulative P-value plots in Figure \ref{fig:figure4}D. There were 166 uncorrected P-values estimated by a serial cut from 125 to 290 for grouping the cohort. All was less than 0.05, and the smallest P-value (\num{8.6e-8}), when cut on n=150 (36.2\% of total cohort 414), was defined as optimal P-value with a cutoff value -0.511(RSEM) of \acrshort{rnaseq}.
%The second candidate, which improving patient survival, is ZNF266.
%\subsubsection{Tables} in main article

%\subsubsection{Figure 4:h}
\begin{figure}[hp]
\centering
\includegraphics[width=\textwidth]{Figure4.pdf}
\bcaption{Kaplan-Meier survival analyses, by cutoff finding.}
{(a) Kaplan-Meier plot of DKK1 under optimal P-value, and (b) the cutoff was derived from cumulative P-value plots of DKK1. (c) Kaplan-Meier plot of ZNF557 under optimal P-value, and (d) the cutoff was derived from cumulative P-value plots of ZNF557.}
\label{fig:figure4}
\end{figure}

>>>>

The 9416 Kaplan-Meier plots with associated Cox's univariate and multivariate tables were generated at workflow step 1 (see Figure \ref{fig:figure1}) and justified by the ranking of hazard ratios.
By uncorrected P-value below 0.05, we selected 967 genes in which \acrfull{hr} is greater than 1.5 or less than 0.5 (see Figure \ref{fig:figure2}(a) univariate, and Figure \ref{fig:figure2}(b)  multivariate plots). At the final step, a Bonferroni P-value correction was used to yield the twenty candidate genes, under the stringent criteria (see Figure \ref{fig:figure2}(c), (d)). The ten candidates, including DKK1, CAMK2N1, STC2, PGK1, SURF4, USP10, NDFIP1, FOXA2, STIP1, and DKC1, have significantly associated with the poor prognosis of \acrshort{os} (see Table \ref{table:table1}),
%\subsubsection{Figure 3}
while the other ten genes were over-expressed in the better survival patients, named as ZNF557, ZNF266, IL19, MYO1H, FCGBP, LOC148709, EVPLL, PNMA5, IQCN (previous name as KIAA1683), and NPB (see Table \ref{table:table3}, with their full name of genes).
We made a volcano plot for 9416 genes by Kaplan-Meier P-value (less than 0.05, obtained during cutoff finding procedure) against the Cox hazard ratio (see Figure \ref{fig:figure3}). The plot revealed that the most significant (Bonferroni-adjusted P-value $< 0.05$) candidate genes are located above the dotted line. At the same time, Cox's HR separated them on the two-side with prognostic impact.

%\subsubsection{Figure 2:h}
\begin{figure}[hp]
\centering
\includegraphics[width=14cm]{Figure2.pdf}
\bcaption{\acrshort{hnscc} Cox's hazard ratio and P-value plots.}
{(a) Univariate HR versus uncorrected P-value; (b) Multivariate HR versus uncorrected P-value; (c) Univariate HR versus Bonferroni corrected P-value; and (d) Multivariate HR versus Bonferroni corrected P-value.}
\label{fig:figure2}
\end{figure}

@@

%\subsubsection{Figure 3:h}
\begin{figure}[hp]
\centering
\includegraphics[width=14cm]{TCGA_HNSC_Optimal_Overall_allPlot_unKM_P_multiHR-Figure3.pdf} % .png from PvalueplotKM_20genes_Bonf.pdf
\bcaption{Volcano plot of genes under survival analyses.}
{X axis: unadjusted P-value of Kaplan-Meier survival (-log10 transformed).
Y axis: multivariate hazard ratio from Cox proportional regression model.
Dotted line: significant Bonferroni corrected P-value. 
Red circles mark 10 candidate genes, which impact on poor prognosis ($HR>=1.5$). Green circles mark 10 genes, which affect on better survival ($HR<=0.5$).}
\label{fig:figure3}
\end{figure}
% new figure of volcano plot: TCGA_HNSC_Optimal_Overall_allPlot_unKM_P_multiHR.pdf, -log10(raw KM P-value) vs Cox multivariate HR
% X axis: "frequency" (number of uncorrected P-values less than 0.05 during Kaplan-Meier cutoff finding procedure). X axis: adjusted Kaplan-Meier P-value.

@@

Our top 1 candidate is \acrfull{DKK1} (see Figure \ref{fig:figure4}(a)). The Kaplan-Meier curve revealed 227 patients bearing the higher expression of \acrshort{DKK1} were suffered from only 40\% of 5-year \acrshort{os} rate. In comparison, the other 187 patients with lower expression (the cutoff at -0.312(RSEM)) have been a better prognosis (adjusted P-value as $0.001$).
Figure \ref{fig:figure4}(b)'s cumulative P-value plot showed that the uncorrected 116 P-values ($< 0.05$) were estimated by a serial cut from 125 to 290 persons for grouping the cohort in our cutoff finding procedure (cutofFinder\_func.R, see Figure \ref{fig:figure1}, "HNSCC cohort"). The smallest P-value (\num{8.9e-8}), when cut on n=187 (45.2\% of 414), was defined as optimal P-value.
Conversely, the most associated gene with better survival is \acrfull{ZNF557}. In Figure \ref{fig:figure4}(c), a Kaplan-Meier curve revealed 264 patients bearing the higher expression of \acrshort{ZNF557} have 55\% of 5-year OS survival rate (adjusted P-value = $0.001$). The cutoff finding procedure (cutofFinder\_func.R) generated cumulative P-value plots in Figure \ref{fig:figure4}(d). There were 166 uncorrected P-values estimated by a serial cut from 125 to 290 for grouping the cohort. All was less than 0.05, and the smallest P-value (\num{8.6e-8}), when cut on n=150 (36.2\% of total cohort 414), was defined as optimal P-value with a cutoff value -0.511(RSEM) of \acrshort{rnaseq}.
%The second candidate, which improving patient survival, is ZNF266.
%\subsubsection{Tables} in main article

%\subsubsection{Figure 4:h}
\begin{figure}[hp]
\centering
\includegraphics[width=15cm]{Figure4.pdf}
\bcaption{Kaplan-Meier survival analyses, by cutoff finding.}
{(a) Kaplan-Meier plot of DKK1 under optimal P-value, and (b) the cutoff was derived from cumulative P-value plots of DKK1. (c) Kaplan-Meier plot of ZNF557 under optimal P-value, and (d) the cutoff was derived from cumulative P-value plots of ZNF557.}
\label{fig:figure4}
\end{figure}
